---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fm.index

fm.index is an R package providing a fast data structure
([FM Index](https://en.wikipedia.org/wiki/FM-index)) for finding occurrences
of string snippets in large libraries of strings (corpus).

Partial string matching can be ~50-fold faster than simple string
scans for many real world string collections. A given corpus is
converted into a compact in-memory FM index representation that can be
efficiently queried for partial string matches.

fm.index wraps the C++ library [SDSL v3](https://github.com/xxsds/sdsl-lite)
and uses a [Compressed Suffix Array](https://en.wikipedia.org/wiki/Compressed_suffix_array)
based on a [Wavelet Tree](https://en.wikipedia.org/wiki/Wavelet_Tree) of the
[Burrow-Wheeler Transform](https://en.wikipedia.org/wiki/Burrows%E2%80%93Wheeler_transform)
of the given string library.

## Installation

Install the most recent [fm.index CRAN release](https://cran.r-project.org/web/packages/fm.index/)

```{r install_cran, echo=TRUE, eval=FALSE}
install.packages("fm.index")
```

You can install the development version from Github with:

```{r install_github, echo=TRUE, eval=FALSE}
install.packages("remotes")
remotes::install_github("clemenshug/fm.index")
```

## Example

In this example we search a library of state names for all states that contain
the substring "new".

```{r example}
library(fm.index)

data("state")
print(state.name)

index <- fm_index_create(state.name, case_sensitive = FALSE)
hits <- fm_index_locate("new", index)
print(hits)
```

The result is a data frame with three columns. `pattern_index` is the index of
the query pattern, `corpus_index` is the index of the matching string in the
string corpus, and `position` is the starting position of the match in the indexed
string. All indices are 1-based.

In order to extract the matching states we can use the `corpus_index` to subset
the vector of original state names.

```{r}
print(state.name[hits$corpus_index])
```



## Performance

The speedup achieved by using fm.index over simple string scans, for example
using `grepl()` or `stringi::stri_locate()`, depends on the kind of strings
in the library.

### Random strings

For this example, we can generate a million random strings of length 50 and
search them for all occurrences of the letters "ab".

```{r}
library(stringi)
library(microbenchmark)

set.seed(42)
lib_random <- stri_rand_strings(1000000, 50)
idx_random <- fm_index_create(lib_random)

head(fm_index_locate("ab", idx_random))

random_benchmark <- microbenchmark(
  fm.index = fm_index_locate("ab", idx_random),
  stringi = stri_locate_all_coll(lib_random, "ab"),
  times = 3
)
print(random_benchmark)
```

Random strings are hard to compress, so searching using an FM Index only yields
a modest ~`r round(1 / do.call("/", as.list(summary(random_benchmark)$median)))`-fold speedup.

### Real-world text

Real-world text is usually repetitive and much easier to compress than
random strings. In this example, we search each line of a book for all instances
of the word "help".

```{r}
book_path <- tempfile()
download.file("http://aleph.gutenberg.org/1/2/3/7/12370/12370.zip", book_path)
book <- scan(unz(book_path, "12370.txt"), sep = "\n", what = "character")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
set.seed(42)
```

Five random lines from the book.

```{r}
book[sample(length(book), 5)]
```

```{r}
idx_book <- fm_index_create(book)
help_locations <- fm_index_locate("water", idx_book)

nrow(help_locations)
head(help_locations)
```

The number of rows in the resulting data frame tells us that the word "water"
occurs `r nrow(help_locations)` times.

```{r}
book_benchmark <- microbenchmark(
  fm.index = fm_index_locate("water", idx_book),
  stringi = stri_locate_all_coll(book, "water"),
  times = 10
)
print(book_benchmark)
```

The text from this book is encoded very efficiently by the FM Index, resulting
in a ~`r round(1 / do.call("/", as.list(summary(book_benchmark)$median)))`-fold speedup.

## Funding

This work was supported by NIH grants U54-HL127365, U24-DK116204, and
U54-HL127624.

## License

This package is provided under the MIT license. The bundled SDSL library is
licensed under the [BSD license](https://github.com/xxsds/sdsl-lite/blob/9930944f14965c4180e40f7acd5f368fd82a3329/LICENSE).
