---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fm.index

fm.index is an R package providing a fast data structure
([FM Index](https://en.wikipedia.org/wiki/FM-index)) for finding occurrences
of string snippets in large libraries of strings.

fm.index wraps the C++ library [SDSL v3](https://github.com/xxsds/sdsl-lite)
and uses a [Compressed Suffix Array](https://en.wikipedia.org/wiki/Compressed_suffix_array)
based on a [Wavelet Tree](https://en.wikipedia.org/wiki/Wavelet_Tree) of the
[Burrow-Wheeler Transform](https://en.wikipedia.org/wiki/Burrows%E2%80%93Wheeler_transform)
of the given string library.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("clemenshug/fm.index")
```
## Example

In this example we search a library of state names for all states that contain
the substring "new".

```{r example}
library(fm.index)

data("state")
print(state.name)

index <- fm_index_create(state.name, case_sensitive = FALSE)
hits <- fm_index_locate("new", index)
print(hits)
```

The result is a data frame with three columns. `pattern_index` is the index of
the query pattern, `library_index` is the index of the matching string in the
index, and `position` is the starting position of the match in the indexed
string. All indices are 1-based.

In order to extract the matching states we can use the `library_index` to subset
the vector of original state names.

```{r}
print(state.name[hits$library_index])
```

## Performance

The speedup achieved by using fm.index over simple string scans, for example
using `grepl()` or `stringi::stri_locate()`, depends on the kind of strings
in the library.

### Random strings

For this example, we can generate a million random strings of length 50 and
search them for all occurrences of the letters "ab". Random strings are hard to
compress, so searching the FM Index only yields a modest ~3-fold speedup.

```{r}
library(stringi)
library(microbenchmark)

set.seed(42)
lib_random <- stri_rand_strings(1000000, 50)
idx_random <- fm_index_create(lib_random)

head(
  fm_index_locate("ab", idx_random)
)

microbenchmark(
  fm.index = fm_index_locate("ab", idx_random),
  stringi = stri_locate_all_coll(lib_random, "ab"),
  times = 10
)
```

### Real-world text

Real-world text is usually repetitive and much easier to compress than
random strings. In this example, we search each line of a book for all
instances of the word “help”. The text from the book is encoded very efficiently
by the FM Index resulting in a ~50-fold speedup.

```{r}
book_path <- tempfile()
download.file("http://aleph.gutenberg.org/1/2/3/7/12370/12370-8.zip", book_path)
book <- scan(unz(book_path, "12370-8.txt"), sep = "\n", what = "character")
idx_book <- fm_index_create(book)

head(
  fm_index_locate("help", idx_book)
)

microbenchmark(
  fm.index = fm_index_locate("help", idx_book),
  stringi = stri_locate_all_coll(book, "help"),
  times = 10
)
```

## Funding

This work was supported by NIH grants U54-HL127365, U24-DK116204, and
U54-HL127624.

## License

This package is provided under the MIT license. The bundled SDSL library is
licensed under the [BSD license](https://github.com/xxsds/sdsl-lite/blob/9930944f14965c4180e40f7acd5f368fd82a3329/LICENSE).
